The Gillespie Algorithm: Stochastic Modeling of Molecular Reaction Networks

### 2024-11-24 paper.Rmd
updated application and references

# Application

@@

# Define reaction rates
c1 <- (1 - (37 + 2) / 1e6) * 100   # Decay of T2
c2 <- (37 / 1e6) * 100             # Formation of CPD
c3 <- (2 / 1e6) * 100              # Formation of 64-PP

@@

# Run the simulation
max_time <- 1000
simulation <- simulate_gillespie(state, max_time)

>>>>

# Application
    We can simulate the dimerization of DNA pyrimidines, a molecular reaction which can lead to skin cancer. Bipyrimidine photodimerization occurs when UV radiation alters the chemical bonds of two consecutive pyrimidine bases, with two possible outcomes: the formation of the cyclobutene pyrimidine dimer (CPD) if a cycloaddition between the C5-C6 double bond of the pyrimidines occurs, or the formation of pyrimidine (6-4) pyrimidone (64-PP) if  a covalent bond forms between C4 and C6 of the pyrimidines (@MartinezFernandez2022).

  Therefore, this simulation involves 4 species and 3 reactions channels: pyrimidine monomers Thymine 1 (T1), Thymine 2 (T2), and dimer photoproducts CPD and 64-PP.

  We will simulate the dimerization of thymine using propensity based on the formation yield of CPD (~37 per 10^6 normal bases) and 64-PP (~2 per 10^6 normal bases) found by exposing skin cells to a UVB dose of 0.2 J/cm^2 (@Mouret2006). As the amount of thymine bases in DNA varies from person to person, we will simulate thymine photodimerization of 1000 pairs of thymine.

@@

# Define reaction rates
c1 <- (1 - (37 + 2) / 1e6) * 10   # No dimerization occurs
c2 <- (37 / 1e6) * 10             # Formation of CPD
c3 <- (2 / 1e6) * 10              # Formation of 64-PP

@@

# Run the simulation
max_time <- 100
simulation <- simulate_gillespie(state, max_time)
